### 2019-05-26 paper/paper.tex
multiwaymerge

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

Sorting is a fundamental programming problem with many important applications, being very useful to organize data for fast retrieval. While there are well-known sorting algorithms for conventional computers, especial techniques are required to take the most out of real hardware offering features such as cache memory and instruction-level parallelism. ChipSort is a Julia module for SIMD and cache-aware sorting. It implements sorting networks and bitonic merge networks with SIMD instructions, with configurable vector sizes. It also implements Combsort, which lends itself easily to vectorization and achieves good performance in arrays that fit cache memory, with low penalty for random access. This also depends on a parallel in-place matrix transposition implemented in ChipSort. Large arrays are approached with a multi-way Mergesort. The implementation of ChipSort itself is of interest from a programming languages perspective due to its use of metaprogramming, and more specifically Julia's generated functions. This enables tuning the code generation to specific tasks and hardware at runtime, a feat owed to Julia. This article presents the implemented techniques as well as experiments that demonstrate the increased speed and the hardware adaptation capability of the module.

\end{abstract}

\section{Introduction}

Sorting has enjoyed an important place within computer programming topics for a long time. Apart from having great importance for practical applications, it also attracts much attention as a theoretically interesting problem in itself~\cite{DBLP:books/lib/Knuth98a}. While sorting is a well-understood problem in general, with a few classic algorithms available, achieving optimal performance for specific problems and architectures may require a careful algorithm selection rather than just tuning parameters \cite[Part II, Introduction]{DBLP:books/daglib/0023376}.

The running time of a computer program is governed by many factors, starting with algorithmic complexity and the base clock speed of the processor. The observed performance in real computers has been increasingly depending in processor features that are not taken into account by the simplest models, though. These features include cache memory and paralellism at the processor and instruction level. Cache memory has become popular in consumer processors in the past couple of decades, allowing programs to often attain higher performance than would be possible with a simpler memory architecture, as long as memory access patterns exhibit temporal and spatial locality~\cite[Chapter 3]{Drepper07whatevery}. Writing software that can better cope with paralellism has also become necessary to exploit the full potential offered by modern and future processors~\cite{wilson2018}.

The interest in sorting and the great relevance of cache memory and paralellism in modern computing naturally lead to research on sorting algorithms that account for these factors. Processors offering SIMD (single instruction multiple data) instructions and other forms of instruction-level parallelism started to become widely available in the early 2000s. One example is the introduction of SSE instructions in the x86 architecture~\cite[Chapter 1]{kusswurm18}. The current availability of processors with 512 bit registers underscore the need of being able to write software that exploit vector operations, not to mention GPUs.

At the time CPUs with SIMD instructions became widely available there was already a trend to employ specialized hardware like GPUs for scientific applications~\cite{larsen2001fast,DBLP:conf/micro/ThompsonHO02}. Althoug SIMD was used mostly for multimedia applications at first~\cite{CHEN2006509,DBLP:journals/mam/SlingerlandS05}, we can cite some early alternative uses including 3D graphics~\cite{DBLP:conf/pcm/MaY02} and machine learning~\cite{DBLP:conf/europar/StreyB01}.

Apart from earlier work on parallel sorting in other architectures, the late 2000s saw the first practical demonstrations of sorting with modern consumer SIMD chips~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/ChhuganiNLMHCBKD08}. Most of the techniques utilized then are still employed in more recent works~\cite{DBLP:journals/pvldb/BalkesenATO13,DBLP:journals/pvldb/InoueT15}.

This article presents \verb ChipSort, a Julia module that implements SIMD and cache-aware sorting utilizing some of the techniques found in the literature. These include sorting and merging networks, the use of Combsort for medium sized arrays and of multi-way merging for large arrays. These techniques require not only tuning parameters such as vector and array sizes, but also generating specific code when implementing different size networks. \verb ChipSort seeks to generate efficient custom code for each task, relying on Julia metaprogramming capabilities for that.

Following Section~\ref{sec:methods} presents the techniques employed in ChipSort. Section~\ref{sec:experiments} presents experimental results assessing the performance of the module and Section~\ref{sec:conclusion} brings some concluding remarks.

\section{SIMD sorting techniques}
\label{sec:methods}
%
ChipSort offers three high-level functions that can be used in applications as replacements for other sorting functions such as \verb sort  in the Julia standard library. Additional work is required from the user selecting among the three functions and also setting some numeric parameters. The functions are intended to be used with arrays of different sizes, and \verb chipsort_large  offers a good overview of all the techniques implemented in the module. Some of these techniques might also be useful in other applications instead of sorting, one clear example being the in-place matrix transpose. ChipSort does not currently support registers with separate key and payload, what might allow significant optimizations. Only the sorting of simple data types such as 32 and 64-bit integers and floating-point numbers were studied.

This section will present how \verb chipsort_large  works, detailing each step of the process. It mostly follows~\cite{DBLP:journals/pvldb/InoueT15} with a few modifications. This function achieves sorting by first splitting the input array in chunks that are sorted separately with \verb chipsort_medium!  and then merged simultaneously. This inner sort is performed in multiple steps:
\begin{enumerate}
\item sorting small blocks of data with a sorting network;
\item reordering vectors in-place according to a matrix transpose;
\item vectorized Combsort with a limited number of iterations;
\item regular insertion sort until the whole data is sorted.
\end{enumerate}

The remainder of this section will detail the multiple techniques employed in this process.

\subsection{Sorting networks}
%
Sorting networks~\cite[Sec. 5.3.4]{DBLP:books/lib/Knuth98a} are graphs composed by comparator modules, units that take two numbers in, and output them in order: the smallest and the largest at specific outputs. Properly arranging the compatators allows the creation of a procedure that takes a vector of numbers in, and outputs them in sequence. The design of a sorting network can minimize the total number of comparisons, or also take into account that comparisons can be done in parallel, minimizing the number of steps to complete the process.

A sorting network can be represented by a Knuth graph such as Figure~\ref{fig:sorting-network}. The graph represents a network that takes an array of four values $(A_1, A_2, A_3, A_4)$ and outputs the sorted sequence $(S_1, S_2, S_3, S_4)$. Each vertical line in the graph represents a comparator.
\begin{figure}[htb]
\centerline{\includegraphics[width=0.6\linewidth]{fig/sorting-network-4.pdf}}
\caption{A sorting network for 4 elements. The dashed lines delimit each step where operations are independent of each other.}
\label{fig:sorting-network}
\end{figure}

In ChipSort, as in the related work, parallelism is mainly exploited by carrying out the comparisons of the sorting network over vectors of size $V$ contained in SIMD registers, effectively running the network $V$ times simultaneously. Other forms of instruction-level parallelism can also be in play, though. In especial, it may also be possible to parallelize instructions such as \verb min  and \verb max , and also the loading of data into registers. This is often done implicitly by the microarchitecture, and programmers can only try to make sure their code is suitable for this by properly ordering operations and avoiding branching, for instance. Part of this work is expected to be performed by the compiler, and especially LLVM in the case of Julia.

ChipSort supports sorting networks of different sizes, currently only powers of 2, which are predefined as data structures in the file \href{https://github.com/nlw0/ChipSort.jl/blob/d2d049b7413f0073476021fa62fb748803130768/src/sorting-network-parameters.jl}{\tt sorting-network-parameters.jl}. The generated function \href{https://github.com/nlw0/ChipSort.jl/blob/d2d049b7413f0073476021fa62fb748803130768/src/sorting-networks.jl#L13}{\tt sort\_net} works by assigning each element of the sequence at each step to a variable, and calculating values from the next step according to the network specification. The comparisons are performed by the {\tt min} and {\tt max} functions, and the function is generic on the element types. Figure~\ref{fig:sort-net-listing}

\begin{figure}[htb]
\begin{lstlisting}[language = Julia]
4 => (((1,2), (3,4)), ((1,3), (2,4)), ((2,3),))
\end{lstlisting}
\begin{lstlisting}[language = Julia]
input_0_1 = input[1]
input_0_2 = input[2]
input_0_3 = input[3]
input_0_4 = input[4]
input_1_1 = min(input_0_1, input_0_2)
input_1_2 = max(input_0_1, input_0_2)
input_1_3 = min(input_0_3, input_0_4)
input_1_4 = max(input_0_3, input_0_4)
input_2_1 = min(input_1_1, input_1_3)
input_2_3 = max(input_1_1, input_1_3)
input_2_2 = min(input_1_2, input_1_4)
input_2_4 = max(input_1_2, input_1_4)
input_3_1 = input_2_1
input_3_4 = input_2_4
input_3_2 = min(input_2_2, input_2_3)
input_3_3 = max(input_2_2, input_2_3)
return (input_3_1, input_3_2, input_3_3, input_3_4)
\end{lstlisting}
\caption{The parameters from a four-element sorting network consisting of three steps, and the corresponding code generated by {\tt sort\_net}.}
\label{fig:sort-net-listing}
\end{figure}

This function is a good first example of the utilization of Julia's meta-programming features in ChipSort. Multiple implementations of sorting networks are available~\cite{DBLP:journals/pvldb/ChhuganiNLMHCBKD08,sortingnetworksjl,ultrasort}, however they implement networks of different sizes straight as code. In ChipSort they are represented as data structures that guide a generated function that produces equivalent code. While this data structure currently is hard-coded, it can potentially be produced only when necessary by an implementation of an algorithm such as Bose-Nelson, what remains a future plan for the module.

The {\tt sort\_net} generated function expects elements with {\tt max} and {\tt min} methods defined. It can be called with simple data types, {\em e.g.} {\tt sort\_net(4,2,5,3)} works. In practice ChipSort often calls this function with vectors, completely relying on SIMD.jl~\cite{erik_schnetter_2019_2592633}. This module maps a vector in Julia similar to tuples or static arrays straight to LLVM vector types, ensuring SIMD code can be produced when possible and also allowing the flexibility of handling logic vectors that are larger than the actual registers.

\subsection{SIMD vectors transpose}
%
Consider a group of stacked SIMD vectors. Each column in this representation is termed a {\em lane}. The result from {\tt sort\_vecs} is that each lane contains a sorted sequence. This data must be transposed to produce sorted vectors. This is done in ChipSort through {\tt transpose\_vecs}, which is a generated function that supports rectangular matrices, but with dimensions constrained to powers of 2. The generated code consists mostly of calls to {\tt SIMD.shufflevec}. This is another instance of an operation that other projects~\cite{DBLP:journals/pvldb/BalkesenATO13,ultrasort} implement with multiple low-level functions, contemplating each different case, and in ChipSort there is only a single higher-level Julia generated function that produces equivalent code.

Figure~\ref{fig:transpose-vecs} displays an example of 64 random integers loaded into SIMD registers as 4 vectors. After {\tt sort\_vecs} each lane contains a sorted sequence, and {\tt transpose\_vecs} turns the lanes into vectors (registers). The \verb chipsort_small!  function is essentially a sorting network followed by a transpose and a merge procedure.

\begin{figure}[htb]
\centerline{\includegraphics[width=0.99\linewidth]{fig/regs.pdf}}
\caption{Four vectors with four values, the result from a vectorized four-elements sorting network, and the transpose.}
\label{fig:transpose-vecs}
\end{figure}

\subsection{Bitonic Merge Networks}
%
A bitonic sequence starts either increasing or decreasing and contains at most one change of direction. Bitonic merge networks allow the creation of merge networks of different sizes, and they are again implemented in ChipSort in a single generated function, {\tt bitonic\_merge}. Merging two sorted vectors requires first reversing one of them, and the result is stored in two new vectors of the same size containing the first and second halves of the total sorted sequence.

Apart from the use in sorting small sequences and in the multi-way merge tree, bitonic merge networks can also be used to implement a regular merge sort. This technique was investigated in ChipSort only superficially, and while there were no promising results in terms of performance, ChipSort also contains the generated function {\tt bitonic\_merge\_interleaved} to illustrate how metaprogramming can help implementing multiple simultaneous merges with interleaved execution, as done in other projects~\cite{DBLP:journals/pvldb/ChhuganiNLMHCBKD08}.

\subsection{Vectorized Combsort}
%
One of the most peculiar aspects of ChipSort, taken directly from AA-sort~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/InoueT15}, is the use of the Combsort algorithm. Combsort is essentially a generalization of Bubble sort in the same way that Shell sort generalizes insertion sort~\cite{dobosiewicz1980efficient,Lacey:1991:FES:117187.117218,DBLP:books/lib/Knuth98a,DBLP:books/daglib/0023376}. The array is sweeped multiple times, and pairs of values at a given distance from each other are compared and stored in order. The distance is reduced at each pass by a factor of $\frac{4}{3}$. This algorithm lends itself easily to vectorization: performing the comparisons over vectors of size $V$ effectively allows us to run the algorithm over $V$ lists simultaneouly.

In practice, ChipSort performs a number of other operations before and after using Combsort. The {\tt chipsort\_medium!} function consists of the following steps:

\begin{enumerate}
\item Apply a sorting network on $K$ blocks of $J$ vectors of size $V$.
\item Vectorized Combsort until the interval size is 1.
\item Transpose blocks.
\item Transpose data in-place from $K$ blocks of $J$ vectors into $J$ blocks of $K$ vectors.
\item Vectorized Combsort again.
\item Sort blocks again.
\item Insertion sort.
\end{enumerate}

After the first two steps the result is essentially a matrix where the $K\times J$ columns are the vectors, and each row contains an approximately sorted sequence. The transpose steps reshape each row into a new block. Each row is processed independently at first, and the transpose allows elements from distinct groups to interact.

The last step is necessary to ensure the whole array is sorted, not vieved anymore as a matrix with independent rows. Insertion sort also presents a better performance than sticking to Combsort until the end, what would essentially mean using bubble sort once the interval becomes 1. Because the array is now approximately sorted, it is expected at this time that insertion sort does not have to move any out-of-order elements very far from where they are found.

\subsection{In-place matrix transpose}
%
In-place matrix transposition can be attained by moving a number to its destination, then moving the number that was found there, and so on, until a cycle is completed and a new number is selected to be moved~\cite{10.1093/comjnl/2.1.47}. While carrying out this procedure is simple, it is not trivial to find out which are the cycles to tranpose a matrix of any given dimensions, and even finding out the number of cycles turns out to be a difficult problem~\cite[Sec. 1.3.3, Ex. 12]{DBLP:books/lib/Knuth97}.

ChipSort contains the generated function {\tt transpose!} to perform in-place matrix transposition. The cycle seeds are computed at the time of code generation, and at runtime the function only needs to move the data and compute the indices detecting when each cycle finishes. The function is therefore completely generic, supporting any dimensions, but with a very simple runtime code, with the most complex part of the problem being tackled programatically during code generation.

\subsection{Vectorized multi-way merging}
%
Merging multiple arrays simultaneously requires the creation of a {\em merge tree} to keep some intermediate data. Each node in this tree is essentially an iterator traversing the merger of two arrays in the lower level. To employ the vectorized bitonic merge network in this process it is necessary to keep a whole vector at each node of the tree. To produce the next vector of elements one of the two source arrays is selected according to their smallest next element, and a whole vector is pulled from that array. This vector is merged with the intermediate one, and the first half of the resulting sequence can be taken to the next level of the tree, while the second half is kept as the new intermediate data.

Given a set of sorted arrays, once the tree is initialized data is taken from it one vector at a time, producing the final sorted sequence. Each time a new vector is requested the tree is traversed looking for the smallest element between each of the node children, until one of the input arrays is reached. The data from the array is then loaded from memory and merged with the intermediate data, producing new updates moving trought the tree until the root is reached.

The intermediate data from the tree is kept by ChipSort in a contiguous array, treated as a heap. Each node ancestor can be found simply by the Euclidean division of its index by 2. During operation it is assumed this data should fit some level of cache memory, as in the related work~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/ChhuganiNLMHCBKD08,DBLP:journals/pvldb/InoueT15}.

\section{Experiments}
\label{sec:experiments}
%
\section{Conclusion}
\label{sec:conclusion}
%

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

% latexmk -bibtex -pdf paper.tex